FictionEro - Affective Impact of Fiction

### 2025-09-02 analysis/analysis.qmd
updated analysis

---
title: "FictionEro - Affective Impact of Fiction"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

@@

results_table <- function(model, effects="fixed", filter=NULL) {
  if("marginaleffects" %in% class(model)) {
    model |> 
      parameters::parameters() |> 
      as.data.frame() |> 
      select(-Parameter, -SE, -S, z=Statistic) |> 
      insight::format_table() |> 
      parameters::display()
  } else {
    display(parameters::parameters(model, effects=effects, keep=filter))
  }
}

@@

@@

### Distributions

@@

@@

#### Study 1

@@

`r report(ttest1, data = dfsub1)`

#### Study 2 
```{r meanarousal 2}
#| message: false

@@

`r report(ttest2, data = dfsub1)`

@@

*Note*. Models for study 1 only include the conditions which were relevant to the participants sexual orientation (e.g., seeing images of men as a heterosexual woman).

```{r}
make_model_plot <- function(model1, model2 = NULL, 
                            title = "Estimated Marginal Means") {

    pred_df <- rbind(
      modelbased::estimate_means(model1, by = c("Gender", "Relevance", "Condition", "ConditionBelief"), backend = "emmeans")  |>
        filter(Relevance == "Relevant") |>
        select(-Relevance) |>
        mutate(Sample = "Study 1")|>
      as.data.frame(), 
      modelbased::estimate_means(model2, by = c("Gender","Condition", "ConditionBelief"), backend = "emmeans")|>
        mutate(Sample = "Study 2") |>
        as.data.frame())

    plt <- pred_df |>
      ggplot(aes(x = ConditionBelief,y = Proportion, color = Condition,linetype = Sample, group = interaction(Sample, Condition))) +
      geom_line(size = 1) +
      geom_point(size = 2) +
      geom_errorbar(aes(ymin = CI_low, ymax = CI_high),width = 0.1) +
      facet_wrap(~ Gender) +
      scale_color_manual(values=c("AI-Generated"="#2196F3", "Photograph"="#F44336")) +
      theme_minimal(base_size = 14) +
        labs(x = "Whether the original labels were believed to be true",y = "Estimated Proportion",color = "Images presented as:",linetype = "Sample", title = "Enticement") +
      theme(
            axis.text.y = element_text(size = 8),
            strip.placement = "outside",
            strip.background.x = element_rect(fill=c("lightgrey"), color=NA),
            strip.text.x = element_text(size = 10),
            strip.text.y = element_text(size = 10),
            axis.text.x = element_text(face = "bold", size=9, color="black"),
            legend.text = element_text(size = 10),
            legend.position =  )

    return(plt)
}

@@

```{r}
arousal1 <- readRDS("models/ModelArousal_1.rds")
arousal2 <- readRDS("models/ModelArousal_2.rds")
```

tbl_a1 <- tbl_results(arousal1, title = "Fixed Effects: Arousal Study 1")
tbl_a2 <- tbl_results(arousal2, title = "Fixed Effects: Arousal Study 2")

plt_a <- make_model_plot(arousal1, arousal2,
                title = "Arousal for Relevant Stimuli")
```

@@

```{r}
enticement1 <- readRDS("models/ModelEnticement_1.rds")
enticement2 <- readRDS("models/ModelEnticement_2.rds")
```

tbl_e1 <- tbl_results(enticement1, title = "Fixed Effects: Enticement Study 1")
tbl_e2 <- tbl_results(enticement2, title = "Fixed Effects: Enticement Study 2")

plt_e <- make_model_plot(enticement1, enticement2,
                title = "Enticement for Relevant Stimuli")
```

@@

```{r}
valence1 <- readRDS("models/ModelValence_1.rds")
valence2 <- readRDS("models/ModelValence_2.rds")
```

tbl_v1 <- tbl_results(valence1, title = "Fixed Effects: Valence Study 1")
tbl_v2 <- tbl_results(valence2, title = "Fixed Effects: Valence Study 2")

plt_v <- make_model_plot(valence1, valence2,
                title = "Valence for Relevant Stimuli")
```

@@

#### Arousal
```{r}
table_dvour <- function(g) {
  performance::performance_dvour(g) |> 
    arrange(desc(D_vour)) |> 
    gt::gt() |> 
    gt::data_color("D_vour", palette = c("red", "red", "orange", "yellow", "green"), domain = c(0, 1)) |> 
    gt::fmt_auto()
}

a1 <- modelbased::estimate_grouplevel(arousal1)
table_dvour(a1)

a2 <- modelbased::estimate_grouplevel(arousal2)
table_dvour(a2)

@@

```{r}
plot_grouplevel_effects <- function(model, outcome_name = NULL) {

  # Colors
  colors <- c("Participant" = "#9C27B0", "Item" = "#FF9800")

  # Plot
  p <- ggplot(model, aes(x = Median, y = Level)) +
    geom_pointrange(aes(xmin = CI_low, xmax = CI_high, color = Group), size = 0.3) +
    ggside::geom_xsidedensity(aes(group = Group, fill = Group), color = NA, alpha = 0.3) +
    scale_color_manual(values = colors) +
    scale_fill_manual(values = colors) +
    theme_minimal() +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.title.x = element_text(size = 12),
      axis.title.y = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank(),
      legend.position = "bottom"
    ) +
    labs(
      x = ifelse(is.null(outcome_name), "Random intercept (participant/item)", paste0(outcome_name, " random intercept")),
      y = "",
      color = "Group",
      fill = "Group"
    )

  return(p)
}

plot_grouplevel_effects(a1, outcome_name = "Arousal")

>>>>

---
title: "FictionEro - Affective Impact of Fiction"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    toc: true
    code-fold: true
    self-contained: false
execute: 
  cache: true
---

@@

@@

make_combined_plot <- function(model1, model2 = NULL, 
                               title = "Estimated Marginal Means") {

  df1 <- modelbased::estimate_means(
    model1, by = c("Gender", "Relevance", "Condition", "ConditionBelief"),backend = "emmeans") |>
    dplyr::mutate(Sample = "Study 1") |>
    as.data.frame()

  # Study 2: only Relevant
  if (!is.null(model2)) {
    df2 <- modelbased::estimate_means(
      model2, by = c("Gender", "Condition", "ConditionBelief"), backend = "emmeans") |>
      dplyr::mutate(Relevance = "Relevant", Sample = "Study 2") |>
      as.data.frame()
    pred_df <- rbind(df1, df2)
  } else {
    pred_df <- df1
  }

  plt <- pred_df |>
    ggplot(aes(x = ConditionBelief, y = Proportion,color = Condition, linetype = Sample, group = interaction(Sample, Condition)
    )) +
    geom_hline(yintercept = 0, color = "darkgrey") +
    geom_line(size = 1) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = CI_low, ymax = CI_high), width = 0.1) +
    facet_grid(Relevance ~ Gender, switch = "y") +
    scale_color_manual(values = c("AI-Generated" = "#2196F3","Photograph" = "#F44336")) +
    theme_minimal(base_size = 14) +
    labs(title = title, x= "Whether the original labels were believed to be true",color = "Images presented as:", linetype = "Sample"
    ) +
    theme(
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 8),
      strip.placement = "outside",
      strip.background.x = element_rect(fill = "lightgrey", color = NA),
      strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10),
      axis.text.x = element_text(face = "bold", size = 9, color = "black"),
      legend.text = element_text(size = 10),
      legend.position = "bottom"
    )

  plt
}

@@

::: {.panel-tabset}

@@

### Arousal Distributions

@@

@@

```{r meanarousal 2}
#| message: false

@@

**Study 1**: `r report::report(ttest1, data = dfsub1)`

**Study 2**: `r report::report(ttest2, data = dfsub2)`

:::

@@

::: {.panel-tabset}

@@

```{r}
arousal1 <- readRDS("models/ModelArousal_1.rds")
arousal2 <- readRDS("models/ModelArousal_2.rds")

tbl_a1 <- tbl_results(arousal1, title = "Fixed Effects: Arousal Study 1")
tbl_a1
```

```{r}
tbl_a2 <- tbl_results(arousal2, title = "Fixed Effects: Arousal Study 2")
tbl_a2
```

```{r}
plta <- make_combined_plot(arousal1, arousal2, title = "Arousal Estimates for All Stimuli")
```

@@

```{r}
enticement1 <- readRDS("models/ModelEnticement_1.rds")
enticement2 <- readRDS("models/ModelEnticement_2.rds")

tbl_e1 <- tbl_results(enticement1, title = "Fixed Effects: Enticement Study 1")
tbl_e1
```

```{r}
tbl_e2 <- tbl_results(enticement2, title = "Fixed Effects: Enticement Study 2")
tbl_e2
```

```{r}
plt_e <- make_combined_plot(enticement1, enticement2, title = "Enticement Estimates for All Stimuli")
plt_e
```

@@

```{r}
valence1 <- readRDS("models/ModelValence_1.rds")
valence2 <- readRDS("models/ModelValence_2.rds")

tbl_v1 <- tbl_results(valence1, title = "Fixed Effects: Valence Study 1")
tbl_v1
```

```{r}
tbl_v2 <- tbl_results(valence2, title = "Fixed Effects: Valence Study 2")
tbl_v2
```

```{r}
plt_v <- make_combined_plot(valence1, valence2, title = "Valence Estimates for All Stimuli")
plt_v
```

:::

@@

@@
